---
title: "Exploring Databrary with R"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Purpose

This document describes some explorations of Databrary using a set of R functions that interact with the Databrary API.

## Use case: Summarizing the demographics of a dataset

Let's say that I want to examine the demographic distribution of a shared dataset. The `databrary_summarize_volume()` command does this for me. It downloads the spreadsheet for a given volume -- the default is volume 4, Adolph, K. (2013). Crawling and walking infants see the world differently. Databrary. Retrieved March 14, 2018 from http://doi.org/10.17910/B7RP4H -- then creates a boxplot of the age distribution by race or ethnicity.

```{r demographics-plot-default}
source("databrary_summarize_volume.R")
databrary_summarize_volume()
```

Here, it appears that the spreadsheet contains some records where the gender was unspecified. These appear to be "Materials" sessions. A future version of the software should eliminate these from the spreadsheet.

Let's try another example from a different dataset.

Tamis-LeMonda, C. (2013). Language, cognitive, and socio-emotional skills from 9 months until their transition to first grade in U.S. children from African-American, Dominican, Mexican, and Chinese backgrounds. Databrary. Retrieved March 14, 2018 from http://doi.org/10.17910/B7CC74.

*I've commented this out because it seems to time out during the knit process. The command does work when run on its own.*

```{r demographics-plot-cathie}
#databrary_summarize_volume(volume = 8)
```
Again, there appear to be some non-session rows in the spreadsheet.

## Use case: Listing all the videos in a session

Let's imagine that I want to list all of the video files in a given session of a specific volume. First, I need to identify the session IDs for a given volume. We'll use the PLAY pilot volume as an illustration.

Adolph, K., Tamis-LeMonda, C. & Gilmore, R.O. (2017). PLAY Pilot Data Collections. Databrary. Retrieved March 14, 2018 from https://nyu.databrary.org/volume/444

```{r list-session-IDs}
source("databrary_download_containers_records.R")
this.vol <- 444
vol444 <- databrary_download_containers_records(volume = this.vol)
str(vol444)
```

There's a lot of information here. The `containers` field of this list contains the session ID numbers. These can be accessed via `vol444$containers$id`.

```{r show-session-ids}
vol444$containers$id
```
 
 Based on the `containers$date` field, it looks like the first entry is not a real session, so let's list the videos in the second entry, the one with the test date equal to `r vol444$containers$date[2]`. 
 
```{r list-videos-date2}
source("databrary_list_assets.R")
this.slot <- vol444$containers$id[2]
databrary_list_assets(volume = 444, slot = this.slot)
```
 
 It looks like there's one movie file there. A quick visit to volume `r this.vol`, slot `r vol444$containers$id[2]` at `r paste0("https://databrary.org/volume/", this.vol, "/slot/", this.slot, "/-")` shows that this looks right.
 
I could download this video, but it's big. So, let's do that in the next vignette with a smaller video file.

## Use case: Downloading the video from a sesssion

Let's download the stimulus samples used in this study:

Gilmore, R.O. (2014). Four-month-olds' discrimination of optic flow patterns depicting different directions of observer motion. Databrary. Retrieved March 14, 2018 from http://doi.org/10.17910/B7Z593.

Let's see how many sessions there are.

```{r sessions-in-gilmore-2014}
this.vol <- 31
vol.31 <- databrary_download_containers_records(volume = this.vol)
vol.31$containers$id
```

We previously showed that the first container/session doesn't usually contain data, so let's look inside the second container/session.

```{r assets-in-vol-31-session-2}
this.slot <- vol.31$containers$id[2]
(these.assets <- databrary_list_assets(volume = this.vol, slot = this.slot))
```

It looks like there are five videos here. Let's download the first one.

```{r download-videos-in-vol-31-slot-9803}
videos <- these.assets$assets$id
source("databrary_download_asset.R")
databrary_download_asset(slot = this.slot, asset = videos[1])
```
That seems to work. Let's see if we can embed it to check. 

```{r embed-sample-video}
(fl <- list.files(path = ".", pattern = "\\.mp4$"))
```

Yep, there's a video called '`r fl[1]`' there. Here it is embedded:

<video controls>
  <source src="`r fl[1]`" type="video/mp4">
Your browser does not support the video tag.
</video>

Let's see if we can download a set of videos.

```{r download-video-list}
lapply(videos, databrary_download_asset, slot = this.slot)
```

That is pretty satisfying.
